---
title: "Training data for structural perception"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup}
library(tidyverse)
library(magrittr)
library(RsSimulx)
library(data.table)
library(viridis)

DIR_DERIVED = "../DerivedData"
DERIVED_COMPLETE_CC = paste( DIR_DERIVED, "TFPerception_deterministic.csv", sep="/")
DERIVED_COMPLETE_Y = paste( DIR_DERIVED, "TFPerception_stochastic.csv", sep="/")
DERIVED_TRAIN_CC = paste( DIR_DERIVED, "TFPerception_deterministic_train.csv", sep="/")
DERIVED_TRAIN_Y = paste( DIR_DERIVED, "TFPerception_stochastic_train.csv", sep="/")
```

# Models
```{r}
study_t <- c(seq(0,4,0.5), 5:12, 16, 24) # sampling points
study_n <- 1000 # sims per group
```


## 1CMT intravascular
```{r}
model.1cmt.iv <- inlineModel("
[LONGITUDINAL]
input = {V, Cl, adde}
EQUATION:
Cc = pkmodel(V,Cl)

DEFINITION:
y = {distribution=lognormal, prediction=Cc, sd=adde}

[INDIVIDUAL]
input = {V_min, V_max, Cl_min, Cl_max }

DEFINITION:
V = {distribution=uniform, min=V_min, max=V_max}
Cl = {distribution=uniform, min=Cl_min, max=Cl_max}")

adm <- list(time=0, amount=1000) # list(time=seq(0,66,by=12), amount=100)
y <- list(name="y", time=study_t) #seq(18, 80, by=6))
Cc <- list(name="Cc", time=study_t ) #seq(0,100, by=0.5))
V <- list(name="V")
p <- c(V_min=1, V_max=200, Cl_min=0.1, Cl_max=20, adde=0.2)
p <- c(V_min=5, V_max=200, Cl_min=1, Cl_max=15, adde=0.2)
g <- list(size=study_n*3) # oversample because we'll remove extreme profiles 

model.1cmt.iv.res <- simulx(model    = model.1cmt.iv, 
              output    = list(Cc,y),
              parameter = p,
              treatment = adm,
              group     = g,
              settings  = list(seed=123456))

hist( model.1cmt.iv.res$parameter$V)
hist( model.1cmt.iv.res$parameter$Cl)

ggplot( model.1cmt.iv.res$Cc, aes(time, Cc, group=id)) +
  geom_line( alpha=0.2) 

prctilemlx(model.1cmt.iv.res$Cc, number=9, level=90, color="#4682b4" ) 
```

remove uninformative profiles

```{r}
#TODO: clean and combine conditions

df.no_elimination <- model.1cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  filter( time == 2, Cc_std > 0.1) %>% 
  ungroup()

dim(df.no_elimination)

df.not1CMT <- model.1cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  filter( time == max(time), Cc_std < 0.2) %>% 
  ungroup()

dim(df.not1CMT)

matched <- sample( intersect( df.no_elimination$id, df.not1CMT$id),
                   size = study_n,
                   replace = F )

model.1cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  ungroup() %>% 
  ggplot(., aes(time, Cc_std, group=id)) +
  geom_line(alpha=0.05) +
  labs( title="Before")

model.1cmt.iv.res$Cc %<>%
  filter( id %in% matched ) 

model.1cmt.iv.res$y %<>%
  filter( id %in% matched )

model.1cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  ungroup() %>% 
  ggplot(., aes(time, Cc_std, group=id)) +
  geom_line(alpha=0.05) +
  labs( title="After")
```

remove uninformative profiles

```{r eval=F}
df <- model.1cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  filter( time == max(time), Cc_std < 0.2) %>% 
  ungroup() %>% 
  arrange( id, time )%>% 
  sample_n( size=study_n, replace = F)

model.1cmt.iv.res$Cc %<>%
  filter( id %in% df$id ) 

model.1cmt.iv.res$y %<>%
  filter( id %in% df$id )

model.1cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  ungroup() %>% 
  ggplot(., aes(time, Cc_std, group=id)) +
  geom_line(alpha=0.05) 
  scale_y_log10()
```

## 1CMT extravascular
```{r}
model.1cmt.oa <- inlineModel("
[LONGITUDINAL]
input = {V, Cl, ka, adde}
EQUATION:
Cc = pkmodel(V,Cl,ka)

DEFINITION:
y = {distribution=lognormal, prediction=Cc, sd=adde}

[INDIVIDUAL]
input = {V_min, V_max, Cl_min, Cl_max, ka_min, ka_max }

DEFINITION:
V = {distribution=uniform, min=V_min, max=V_max}
Cl = {distribution=uniform, min=Cl_min, max=Cl_max}
ka = {distribution=uniform, min=ka_min, max=ka_max}" )

adm <- list(time=0, amount=1000) # list(time=seq(0,66,by=12), amount=100)
y <- list(name="y", time=study_t) #seq(18, 80, by=6))
Cc <- list(name="Cc", time=study_t ) #seq(0,100, by=0.5))
V <- list(name="V")
Cl <- list(name="Cl")
ka <- list(name="ka")
g <- list(size=study_n*2) # oversample, to remove uninformative profiles

p <- c(V_min=1, V_max=100, Cl_min=0.1, Cl_max=20, ka_min=0.1, ka_max=0.7, adde=0.2)

model.1cmt.oa.res <- simulx(model    = model.1cmt.oa, 
              output    = list(Cc,y, V, Cl, ka),
              parameter = p,
              treatment = adm,
              group     = g,
              settings  = list(seed=42))

ggplot( model.1cmt.oa.res$Cc, aes(time, Cc, group=id)) +
  geom_line( alpha=0.2) 

prctilemlx(model.1cmt.oa.res$Cc, number=9, level=90, color="#4682b4" ) 
```

Remove untypical profiles (less than 20% Cend/Cmax)

```{r}
df <- model.1cmt.oa.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  filter( time == max(time), Cc_std < 0.2) %>% 
  ungroup() %>% 
  arrange( id, time )%>% 
  sample_n( size=study_n, replace = F)

model.1cmt.oa.res$Cc %<>%
  filter( id %in% df$id ) 

model.1cmt.oa.res$y %<>%
  filter( id %in% df$id )

model.1cmt.oa.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  ungroup() %>% 
  ggplot(., aes(time, Cc_std, group=id)) +
  geom_line(alpha=0.05) 
  scale_y_log10()
```


## 2CMT intravascular
```{r}
model.2cmt.iv <- inlineModel("
[LONGITUDINAL]
input = {V, Cl, Q, V2, adde}
EQUATION:
Cc = pkmodel(V, k=Cl/V, k12=Q/V, k21=Q/V2)

DEFINITION:
y = {distribution=lognormal, prediction=Cc, sd=adde}

[INDIVIDUAL]
input = {V_min, V_max, Cl_min, Cl_max, V2_min, V2_max, Q_min, Q_max }

DEFINITION:
V = {distribution=uniform, min=V_min, max=V_max}
Cl = {distribution=uniform, min=Cl_min, max=Cl_max}
V2 = {distribution=uniform, min=V2_min, max=V2_max}
Q = {distribution=uniform, min=Q_min, max=Q_max}")

adm <- list(time=0, amount=1000) # list(time=seq(0,66,by=12), amount=100)
y <- list(name="y", time=study_t) #seq(18, 80, by=6))
Cc <- list(name="Cc", time=study_t ) #seq(0,100, by=0.5))
V <- list(name="V")
p <- c(V_min=10, V_max=100, Cl_min=2, Cl_max=10, 
       V2_min=10, V2_max=1000, Q_min=1, Q_max=20, adde=0.2)
g <- list(size=study_n*2) 

model.2cmt.iv.res <- simulx(model    = model.2cmt.iv, 
              output    = list(Cc,y),
              parameter = p,
              treatment = adm,
              group     = g,
              settings  = list(seed=123456))

ggplot( model.2cmt.iv.res$Cc, aes(time, Cc, group=id)) +
  geom_line( alpha=0.2) 

prctilemlx(model.2cmt.iv.res$Cc, number=9, level=90, color="#4682b4" ) 
```

remove uninformative profiles

```{r}
#TODO: clean and combine conditions

df.no_elimination <- model.2cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  filter( time == 2, Cc_std > 0.2) %>% 
  ungroup()

dim(df.no_elimination)

df.not2CMT <- model.2cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  filter( time == 6, Cc_std < 0.4) %>% 
  ungroup()

matched <- sample( intersect( df.no_elimination$id, df.not2CMT$id),
                   size = study_n,
                   replace = F )

model.2cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  ungroup() %>% 
  ggplot(., aes(time, Cc_std, group=id)) +
  geom_line(alpha=0.05) +
  labs( title="Before")

model.2cmt.iv.res$Cc %<>%
  filter( id %in% matched ) 

model.2cmt.iv.res$y %<>%
  filter( id %in% matched )

model.2cmt.iv.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  ungroup() %>% 
  ggplot(., aes(time, Cc_std, group=id)) +
  geom_line(alpha=0.05) +
  labs( title="After")
```

## 2CMT extravascular
```{r}
model.2cmt.oa <- inlineModel("
[LONGITUDINAL]
input = {V, Cl, Q, V2, ka, adde}
EQUATION:
Cc = pkmodel(V, k=Cl/V, k12=Q/V, k21=Q/V2, ka)

DEFINITION:
y = {distribution=lognormal, prediction=Cc, sd=adde}

[INDIVIDUAL]
input = {V_min, V_max, Cl_min, Cl_max, V2_min, V2_max, Q_min, Q_max, ka_min, ka_max }

DEFINITION:
V = {distribution=uniform, min=V_min, max=V_max}
Cl = {distribution=uniform, min=Cl_min, max=Cl_max}
V2 = {distribution=uniform, min=V2_min, max=V2_max}
Q = {distribution=uniform, min=Q_min, max=Q_max}
ka = {distribution=uniform, min=ka_min, max=ka_max}")

adm <- list(time=0, amount=1000) # list(time=seq(0,66,by=12), amount=100)
y <- list(name="y", time=study_t) #seq(18, 80, by=6))
Cc <- list(name="Cc", time=study_t ) #seq(0,100, by=0.5))
V <- list(name="V")
p <- c(V_min=1, V_max=100, Cl_min=0.1, Cl_max=50, 
       V2_min=10, V2_max=1000, Q_min=1, Q_max=50, 
       ka_min=0.1, ka_max=0.7, adde=0.2)
g <- list(size=study_n*2) 

model.2cmt.oa.res <- simulx(model    = model.2cmt.oa, 
              output    = list(Cc,y),
              parameter = p,
              treatment = adm,
              group     = g,
              settings  = list(seed=123456))

df <- model.2cmt.oa.res$Cc %>% 
  group_by(id) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc))) %>% 
  filter( time == max(time), Cc_std < 0.2) %>% 
  ungroup() %>% 
  arrange( id, time )%>% 
  sample_n( size=study_n, replace = F)

model.2cmt.oa.res$Cc %<>%
  filter( id %in% df$id ) 

model.2cmt.oa.res$y %<>%
  filter( id %in% df$id )

ggplot( model.2cmt.oa.res$Cc, aes(time, Cc, group=id)) +
  geom_line( alpha=0.2) 

prctilemlx(model.2cmt.oa.res$Cc, number=9, level=90, color="#4682b4" ) 
```
# Datasets
Generating the complete training datasets

## Store
### Cc
Deterministic sims
```{r}
df.Cc <- rbind( 
  model.1cmt.iv.res$Cc %>% 
    mutate( outcome = "1CMT_IV" ),
  model.2cmt.iv.res$Cc %>% 
    mutate( outcome = "2CMT_IV" ),
  model.1cmt.oa.res$Cc %>% 
    mutate( outcome = "1CMT_OA" ),
  model.2cmt.oa.res$Cc %>% 
    mutate( outcome = "2CMT_OA" ) ) 

ggplot( df.Cc, aes(time, Cc, group=interaction(id, outcome)) ) +
  geom_line( aes(color=outcome), alpha=0.2) +
  labs( title="Training data (deterministic)",
        x="Time",
        y="Concentration",
        color="Type") +
  theme_minimal()

df.Cc %>% 
  select( id, time, DV=Cc, outcome) %>% 
  write.csv( ., DERIVED_COMPLETE_CC, row.names = F)
```

### y
Stochastic sims
```{r}
df.y <- rbind( 
  model.1cmt.iv.res$y %>% 
    mutate( outcome = "1CMT_IV" ),
  model.2cmt.iv.res$y %>% 
    mutate( outcome = "2CMT_IV" ),
  model.1cmt.oa.res$y %>% 
    mutate( outcome = "1CMT_OA" ),
  model.2cmt.oa.res$y %>% 
    mutate( outcome = "2CMT_OA" ) ) 

ggplot( df.y, aes(time, y, group=interaction(id, outcome)) ) +
  geom_line( aes(color=outcome), alpha=0.2) +
  labs( title="Training data (stochastic)",
        x="Time",
        y="Concentration",
        color="Type") +
  theme_minimal()

df.y %>% 
  select( id, time, DV=y, outcome) %>% 
  write.csv( ., DERIVED_COMPLETE_Y, row.names = F)
```

## Training data

```{r}
study_timepoints <- c(0, 1, 2, 4, 6, 8, 12, 24 ) 
```


### Deterministic

```{r}
df.Cc %<>% 
  group_by(id, outcome) %>%  
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc)),
          id = as.numeric(levels(id))[id] )

ggplot( df.Cc, aes(time, Cc_std, color=id, group=id)) +
  geom_line( alpha=0.05) +
  scale_color_viridis( option="inferno" )+
  theme_minimal() +
  theme( legend.position = "none") +
  facet_wrap( .~outcome) +
  labs( title="Training Data",
        subtitle=paste( "Simulations per model:", study_n),
        y="") 
```

Select sampling time points, pivot from long to wide, and store. Re-run standardization on this subset so as to prevent cross-over.

```{r}
df <- df.Cc %>% 
  filter( time %in% study_timepoints) %>% 
  group_by( id, outcome ) %>% 
  mutate( Cc_std = (Cc- min(Cc)) / (max(Cc) - min(Cc)) ) %>% 
  select( id, outcome, Cc_std, time ) %>% 
  pivot_wider( names_from = "time", names_prefix="t", values_from = "Cc_std")

write.csv( df, DERIVED_TRAIN_CC, quote = F, row.names = F)
```

### Stochastic
Same thing for the y data

```{r}
df.y %<>% 
  group_by(id, outcome) %>%  
  mutate( y_std = (y- min(y)) / (max(y) - min(y)),
          id = as.numeric(levels(id))[id] )

ggplot( df.y, aes(time, y_std, color=id, group=id)) +
  geom_line( alpha=0.05) +
  scale_color_viridis( option="inferno" )+
  theme_minimal() +
  theme( legend.position = "none") +
  facet_wrap( .~outcome) +
  labs( title="Training Data (stochastic)",
        subtitle=paste( "Simulations per model:", study_n),
        y="") 
```

And selecting our time points

```{r}
df <- df.y %>% 
  filter( time %in% study_timepoints ) %>% 
  group_by( id, outcome ) %>% 
  mutate( y_std = (y- min(y)) / (max(y) - min(y)) ) %>% 
  select( id, outcome, y_std, time ) %>% 
  pivot_wider( names_from = "time", names_prefix="t", values_from = "y_std")

write.csv( df, DERIVED_TRAIN_Y, quote = F, row.names = F)
```


